### 2024-04-15 paper/result_table.tex
Added Performance report prototype latex

% !TEX root = paper.tex
\documentclass{article}
\usepackage[margin=1in]{geometry} % Adjust document margins
\usepackage{lipsum} % Dummy text
\usepackage{multicol} % Multi-column layout for the body
\usepackage[hidelinks]{hyperref}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{tikz}
\usepackage{float}
\usepackage{booktabs}
\usepackage{amsthm}
\usepackage{multirow}
\usepackage [english]{babel}
\usepackage [autostyle, english = american]{csquotes}

\graphicspath{ {../images/} }

\title{Knowledge and Story Comprehension}
\author{Eduardo Badillo\thanks{UC Berkeley, Email: eduardo.badillo@berkeley.edu} \and Geronimo Walker\thanks{UC Berkeley, Email: 
geronimowalker@ischool.berkeley.edu} \and   Svein Gonzalez\thanks{UC Berkeley, Email: 
sggonzal@ischool.berkeley.edu}}
\date{\today}

\begin{document}

\maketitle

\begin{table}[h!]
    \centering
    \begin{tabular}{lcccc}
    \hline
    \multirow{2}{*}{Model} & \multicolumn{2}{c}{Alternate Story Endings} & \multicolumn{2}{c}{GPT Story Endings} \\
    & Accuracy & Recall & Accuracy & Recall \\
    \hline
    Baseline Inference & 82.5 & NA & 68.55 & NA \\
    Inference w/ knowledge & 90.38 & NA & 61.92 & NA \\
    Triple Inference & 71.63 & NA & 66.80 & NA \\
    \hline
    Fine-tune Seq2Seq & 96.96 & 97.11 & 49.41 & 100 \\
    Knowledge Seq2Seq & NA & NA & 50.15 & 100 \\
    Triple Seq2Seq & 94.50 & 94.21 & 80.33 & 81.82 \\

    \end{tabular}
    \caption{Model performance for Alternate Story Endings and GPT Story Endings.}
    \label{tab:model_performance}
    \end{table}

\end{document}